Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-03-30 fib2.tex
Trying out having a two-column figure just to see how it looks. Thoughts?
minor text edits

A physical anyon model is abstractly described by a unitary modular tensor category~\cite{Wang2010b}. This object contains the data describing the types of anyonic particles, as well as the results of topological operations such as braiding or fusion of these anyons. The Fibonacci anyon model consists of only one non-trivial particle type, conventionally called $\tau$. Denoting the vacuum by $\mathbb{I}$, we can describe the possible fusion outcomes of Fibonacci anyons as $\tau\times\tau=\mathbb{I}+\tau$, i.e.~two $\tau$ particles can either fuse to vacuum or to a $\tau$ particle.

Associated with the fusion outcomes of a fixed number of $\tau$ particles is a Hilbert space, the \emph{fusion} space. For $n$ particles of type $\tau$, the dimension of this space grows asymptotically as $\varphi^n$ for $\varphi=\frac{1+\sqrt{5}}{2}$ the golden ratio. Additionally, there is a global degeneracy associated with the topology of the manifold on which the anyons reside. We will consider systems with the topology of a torus, which for the Fibonacci anyons gives rise to a 2-fold degeneracy. For convenience (in particular to minimize finite-size effects in our numerical computations~\cite{Brell2013}), this global space is the one in which we will encode, and demonstrate protection of, quantum information.

The fusion outcome of any particular experiment depends on the history of the particles, in particular how they have braided around one another. For this reason braiding operations are considered to act unitarily on the fusion space. Pair-creation events should be considered isometries that enlarge the fusion space, while fusion events correspond to projective 
measurements of combined charge of some number of particles. For details of the description of these processes in Fibonacci anyons, see Ref.~\cite{Trebst2008}. \stf{check that this ref is a good one...} Braiding of Fibonacci anyons and measurements of fusion outcomes is known to be sufficient to implement universal quantum computation~\cite{?}.

We use a phenomenological model of Fibonacci anyon dynamics which neglects any microscopic details of the system. This is consistent with the principles of topologically ordered systems and anyonic physics, where the key universal features describing the anyon model correspond to large length-scale physics, while the microscopic physics plays a less 
important (and non-universal) role. As such, we model the system simply as a square lattice of sites on which sets of anyons can reside. Allowed anyon dynamics are considered to be local on this lattice, which we take to have periodic boundary conditions. For more details of an analogous model, see~\cite{Brell2013}.

@@

	We consider encoding a qubit of quantum information in the global topological degeneracy associated with the topology of the manifold of our system. As such, the codespace is specified as having vacuum charge at each site of the lattice. Typical realistic noise processes in this kind of system are pair-creation, hopping, exchange etc.~of anyons on neighbouring sites. It was seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only is sufficient to capture the qualitative features of an error-correction simulation, and so we will restrict to pair-creation noise processes in this study for convenience. These events act on neighbouring pairs of sites, and so may be associated with an edge of the lattice, chosen uniformly at random at each timestep of the simulation. The simulation time, and thus the error-correction threshold, will be measured in terms of average number of noise processes per edge, as opposed to an iid noise probability per edge. The latter measure is not appropriate for error processes in nonabelian anyon models, but the two coincide in the limit of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

	In order to perform a logical error on our code, a noise process must have support on a ribbon running around a homologically nontrivial loop of the lattice. These correspond to processes in which a $\tau$ anyon is transported around the non-trivial loop before annihilating to vacuum. More details can be found in Ref.~\cite{?}.

@@

@@

\begin{figure}[th!]
\begin{center}
	{\scriptsize (i)
	\includegraphics[width=0.27\columnwidth]{pair-create.pdf}
	\hskip 4pt
	(ii)
	\includegraphics[width=0.27\columnwidth]{syndrome-1.pdf}
	\hskip 4pt
	(iii)}
	\includegraphics[width=0.27\columnwidth]{syndrome-2.pdf}
\caption{The simulation of noise processes proceeds by i) creating pairs of anyons on neighbouring tiles represented by open curves ii) joining the curves within a given tile iii) measuring the total charge on each tile.}
\label{f:simex}
\end{center}
\end{figure}

>>>>

The topological features of a physical anyon model are abstractly described by a unitary modular tensor category~\cite{Wang2010b}. This object contains the data describing the types of anyonic particles, as well as the results of topological operations such as braiding or fusion of these anyons. The Fibonacci anyon model consists of only one non-trivial particle type, conventionally labelled $\tau$. Denoting the vacuum by $\mathbb{I}$, we can describe the possible fusion outcomes of Fibonacci anyons as $\tau\times\tau=\mathbb{I}+\tau$, i.e.~two $\tau$ particles can either fuse to vacuum or to a $\tau$ particle.

Associated with each set of fusion outcomes for a fixed number of $\tau$ particles is a basis vector for a Hilbert space, the \emph{fusion} space. For $n$ particles of type $\tau$, the dimension of this space grows asymptotically as $\varphi^n$ for $\varphi=\frac{1+\sqrt{5}}{2}$ the golden ratio. Additionally, there is a global degeneracy associated with the topology of the manifold on which the anyons reside. We will consider systems with the topology of a torus, which for the Fibonacci anyons gives rise to a 2-fold degeneracy. For convenience (in particular to minimize finite-size effects in our numerical computations~\cite{Brell2013}), this global space is the one in which we will encode, and demonstrate protection of, quantum information.

The fusion outcome of any particular experiment depends on the history of the particles, in particular how they have braided around one another. Braiding operations are treated as unitary actions on the fusion space. Pair-creation events should be considered isometries that enlarge the fusion space, while fusion events correspond to projective 
measurements of combined charge of some number of particles. For details of the description of these processes in Fibonacci anyons, see Ref.~\cite{Trebst2008}. \stf{check that this ref is a good one...} Braiding of Fibonacci anyons and measurements of fusion outcomes is known to be sufficient to implement universal quantum computation~\cite{?}.

We use a phenomenological model of Fibonacci anyon dynamics which neglects any microscopic details of the system. This is consistent with the principles of topologically ordered systems and anyonic physics, where the key universal features describing the anyon model correspond to large length-scale physics, while the microscopic physics plays a less 
important (and non-universal) role. As such, we model the system simply as a periodic $L\times L$ square lattice $\Lambda$ of sites with upon which sets of anyons can reside. Allowed anyon dynamics (pair-creation, braiding, and fusion) are considered to act locally on edges of $\Lambda$. For more details of an analogous model, see~\cite{Brell2013}.

@@

	We consider encoding a qubit of quantum information in the global degeneracy associated with the topology of the manifold of our system. We imagine an idealized Hamiltonian for this system of the form
	\begin{align}
		H=-\sum_{i\in Lambda}\proj{\mathbb{I}_i}\;,\label{e:hamiltonian}
	\end{align}
with $\proj{\mathbb{I}}_i$ the projector to charge $\mathmm{I}$ at site $i$, i.e.~the ground space of the model has vacuum (or no anyon) at each site. Given the toroidal boundary conditions, this space is two-fold degenerate as noted above. We identify this ground space as the codespace of our model. Typical realistic noise processes in this kind of system are pair-creation, hopping, exchange etc.~of anyons on neighbouring sites of $\Lambda$. It was seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only is sufficient to capture the qualitative features of an error-correction simulation, and so we will restrict to pair-creation noise processes in this study for convenience. These events act on neighbouring pairs of sites, and so may be associated with an edge of the lattice, chosen uniformly at random at each timestep of the simulation. The simulation time, and thus the error-correction threshold, will be measured in terms of average number of noise processes per edge, as opposed to an iid noise probability per edge. The latter measure is not appropriate for error processes in nonabelian anyon models, but the two coincide in the limit of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

	In order to perform a logical error on our code, a noise process must have support on a ribbon running around a homologically nontrivial loop of the lattice. These correspond to processes in which anyonic charge is transported around the non-trivial loop before annihilating to vacuum. More details can be found in Ref.~\cite{?}.

@@

\begin{figure*}[th!]
\begin{center}
	{\scriptsize (i)
	\includegraphics[width=0.57\columnwidth]{pair-create.pdf}
	\hskip 4pt
	(ii)
	\includegraphics[width=0.57\columnwidth]{syndrome-1.pdf}
	\hskip 4pt
	(iii)}
	\includegraphics[width=0.57\columnwidth]{syndrome-2.pdf}
\caption{The simulation of noise processes proceeds by i) creating pairs of anyons on neighbouring tiles represented by open curves ii) joining the curves within a given tile iii) measuring the total charge on each tile.}
\label{f:simex}
\end{center}
\end{figure*}

@@
